Improving reproducibility of scientific software using Nix/NixOS: A case study on the preCICE ecosystem

### 2024-05-14 paper/paper.tex
fix(paper): Trailing whitespace
feat(paper): Resolve some todos

We present challenges, solutions, and workarounds for all software packages of the preCICE ecosystem.
\todo{mention where and which results to find on GitHub}

@@

\todo{matlab and fortran module missing}

@@

The \texttt{pyprecice} package is already available upstream, but it is currently broken: It does not compile as it is not able to find the python module \texttt{pkgconfig}.
After adding this single dependency as an input to the package definition, the python package builds and can be used as expected.

\todo{fixed upstream?}

\subsubsection{Julia-bindings}

@@

After successfully packaging all dependencies, the build phase of Code\_Aster completes, but the installation part has another issue, which also affects new versions of Ubuntu and other distributions using a python version greater than 3.9.
There is a forum post without resolution, so we need to manually patch the bug. \todo{link post}
The issue is precisely, that the custom build system does not correctly calculate the python site-packages directory. It unconditionally slices the first 3 chars from \texttt{sys.version}, which works for version 3.9.x, but not for version 3.10.x.
After patching, Code\_Aster successfully installed into \texttt{\$out/14.6/}, so we move around some files until we have a valid directory structure with \texttt{\$out/bin}, \texttt{\$out/lib}, providing symlinks for \texttt{\$out/14.6/} and \texttt{\$out/stable/}.

@@

The Nix code for the packages for the case study on the preCICE adapters, solvers and the preCICE VM is available online in a git repository at \url{https://github.com/precice/nix-packages/} with a more in-detail report that also includes a comparison between Nix, Spack and EasyBuild and a evaluation of Nix on HPC~\footnote{\url{https://github.com/precice/nix-packages/releases/tag/initial-paper-release}}.

>>>>

We present challenges, solutions, and workarounds for all software packages of the preCICE ecosystem.
The Nix code for the packages for the case study on the preCICE adapters, solvers and the preCICE VM is available online in a git repository at \url{https://github.com/precice/nix-packages/} with a more in-detail report that also includes a comparison between Nix, Spack and EasyBuild and a evaluation of Nix on HPC~\footnote{\url{https://github.com/precice/nix-packages/releases/tag/initial-paper-release}}.

@@

\subsubsection{Matlab bindings}

There were several attempts to package matlab\footnote{\url{https://de.mathworks.com/products/matlab.html}} for NixOS in the past\footnote{\url{https://github.com/NixOS/nixpkgs/issues/56887}}, yet there was no success so far.
This might be due to the fact, that matlab needs to be installed by running an installation wizard that downloads files during the installation process and verifies the licence.
Also, as matlab is not installed in the VM, we decided to not proceed in packaging the software.

\subsubsection{Fortran bindings}

preCICE provides bindings\footnote{\url{https://github.com/precice/fortran-module}} to the fortran programming language.
We were able to package these by providing custom phases for the \texttt{buildPhase} and the \texttt{installPhase}.
The resulting binary could run the example that comes with the module.

@@

The \texttt{pyprecice} package is already available upstream, but did not compile as it is not able to find the python module \texttt{pkgconfig}.
After adding this single dependency as an input to the package definition, the python package builds and can be used as expected.
This solution was contributed to the upstream nixpkgs repository.

\subsubsection{Julia bindings}

@@

After successfully packaging all dependencies, the build phase of Code\_Aster completes, but the installation part has another issue, which also affects new versions of Ubuntu and other distributions using a python version greater than 3.9.
There is a forum post\footnote{\url{https://forum.code-aster.org/public/d/26475-problem-installing-code-aster-version-14-6/11}} without resolution, so we need to manually patch the bug.
The issue is precisely, that the custom build system does not correctly calculate the python site-packages directory. It unconditionally slices the first 3 chars from \texttt{sys.version}, which works for version 3.9.x, but not for version 3.10.x.
After patching, Code\_Aster successfully installed into \texttt{\$out/14.6/}, so we move around some files until we have a valid directory structure with \texttt{\$out/bin}, \texttt{\$out/lib}, providing symlinks for \texttt{\$out/14.6/} and \texttt{\$out/stable/}.

@@
